\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-15 paper/main.tex
Add explanation for AUPRC, some citations
Finished abstract, updated README

\abstract{\textbf{Motivation:} Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text
Text Text Text Text Text.\\
\textbf{Results:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text\\
\textbf{Availability:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text  Text\\
\textbf{Contact:} \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}\\
\textbf{Supplementary information:}10264703 Supplementary data are available at \textit{Bioinformatics}
online.}

@@

We developed \name{} as a method for predicting DTIs. We use an
ontology-based machine learning method \cite{DL2vec2020} to encode phenotypic
consequences of DTIs and deep learning methods to encode molecular
features. We combine both using a protein interaction network which we
exploit using a graph neural network. We use this model to test
whether molecular or phenotype features benefit from the network
information and find that only phenotype features localize on the
graph whereas molecular features do not. We further evaluate and
compare \name{} against several DTI prediction methods and
demonstrate a substantial improvement of \name{} over the state of the
art in predicting drugs that target a protein. We also identify and
characterize several biases in both training and evaluating DTI
prediction methods, and make recommendations on how to avoid
them. \name{} is available as Free Software at \url{}\todo{add}.

@@

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. For those entities, we impute the $MicroAUC$
interpolating linearly, by using the accuracy for this
subset:
\todo[inline]{Is the accuracy usually close to $1$? I am still not so
  sure this imputation with accuracy is well justified -- is there a
  precedent/paper to cite somewhere?}
\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).

>>>>

\abstract{\textbf{Motivation:} \textit{In silico} drug--target interaction (DTI) prediction is crucial for drug discovery and down-stream tasks, such as drug indication and drug--disease prediction. However, DTI datasets are prone to certain biases, skewing the effective predictive expressiveness over those datasets. Additionally, only few approaches combine phenotypical and molecular features for DTI prediction.\\
\textbf{Results:} We present \name{} to combine molecular features and functional information with protein--protein interaction networks under usage of graph convolutional neural networks, showing that phenotypic information localizes on the interaction network. We hereby focus on protein-centric evaluation, predicting drugs that may target specific targets. We further show that common drug--target interaction datasets contain intrinsic biases, majorly affecting predictive performance of compared models. Additionally, we propose an modified evaluation scheme and metric to circumvent such skews and to better reflect the problem DTI prediction methods aim to solve. \\
\textbf{Availability:} Source code and data used in \name{} are available at \url{https://github.com/THinnerichs/DTI-VOODOO}\\
\textbf{Contact:} \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}\\
\textbf{Supplementary information:}10264703 Supplementary data are available at \textit{Bioinformatics}
online.}

@@

We developed \name{} as a method for predicting DTIs. We use an
ontology-based machine learning method \cite{DL2vec2020} to encode phenotypic
consequences of DTIs and deep learning methods to encode molecular
features. We combine both using a protein interaction network which we
exploit using a graph neural network. We use this model to test
whether molecular or phenotype features benefit from the network
information and find that only phenotype features localize on the
graph whereas molecular features do not. We further evaluate and
compare \name{} against several DTI prediction methods and
demonstrate a substantial improvement of \name{} over the state of the
art in predicting drugs that target a protein. We also identify and
characterize several biases in both training and evaluating DTI
prediction methods, and make recommendations on how to avoid
them. \name{} is available as Free Software at \url{https://github.com/THinnerichs/DTI-VOODOO}.

@@

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. For those entities, we impute the $MicroAUC$
interpolating linearly, by using the accuracy for this
subset:

\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).

Further, we choose AUROC over area under precision recall curve (AUPRC) as the latter is prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases, undermining cross-approach evaluation.\todo[inline]{More Information? Right place?}
